---
title: "p8105_hw3_sy2824"
author: "Sijia Yue"
date: "2018-10-06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)

```

## Cleaning data

```{r}
tidy_brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location))  %>% 
  mutate(response = forcats::fct_relevel(response,c("Excellent", "Very good", "Good", "Fair", "Poor")))
```

## Questions

**In 2002, which states were observed at 7 locations?**
```{r}
filter(tidy_brfss, year == "2002") %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  filter(n_location == "7") 
```
In 2002, the states that were observed at 7 locations are `CT`, `FL` and `NC`.

**Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**
```{r}
  group_by(tidy_brfss,locationabbr,year) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  ggplot(aes(y = n_location, x = year)) +
  geom_line(aes(color = locationabbr), alpha = .5) +
  labs(
    title = "Spaghetti plot",
    x = "Year",
    y = "Number of observations in each state"
  ) 
  
```

**Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**
```{r}
filter(tidy_brfss, year == "2002" | year == "2006" | year == "2010") %>% 
  filter(locationabbr == "NY") %>% 
  spread( key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(sum = excellent + very_good + good + fair + poor) %>% 
  mutate(excellent_proportion = (excellent)/sum) %>% 
  group_by(year) %>% 
  summarize(
    mean_proportion = mean(excellent_proportion),
    sd_proportion = sd(excellent_proportion)
    ) 
```

**For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**
```{r}
spread(tidy_brfss, key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(sum = excellent + very_good + good + fair + poor) %>% 
  mutate(excellent_proportion = excellent / sum) %>% 
  mutate(very_good_proportion = very_good / sum) %>% 
  mutate(good_proportion = good / sum) %>% 
  mutate(fair_proportion = fair / sum) %>% 
  mutate(poor_proportion = poor / sum) %>% 
  group_by(year, locationabbr) %>% 
  summarize(mean_excellent_proportion = mean(excellent_proportion),
            mean_very_good_proportion = mean(very_good),
            mean_good_proportion = mean(good_proportion),
            mean_fair_proportion = mean(fair_proportion),
            mean_poor_proportion = mean(poor_proportion))
  
```

```{r}
tidy_brfss %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(mean = mean(data_value, na.rm = TRUE))
```

