---
title: "p8105_hw3_sy2824"
author: "Sijia Yue"
date: "2018-10-06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  fig.width = 9,
  fig.height = 9,
  fig.asp = .9,
  out.width = "100%"
)
library(tidyverse)
library(p8105.datasets)
library(rnoaa)
library(patchwork)

```

# Question 1
## Cleaning data

```{r}
tidy_brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location))  %>% 
  mutate(response = forcats::fct_relevel(response,c("Excellent", "Very good", "Good", "Fair", "Poor")))
```

## Questions

**In 2002, which states were observed at 7 locations?**
```{r}
filter(tidy_brfss, year == "2002") %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  filter(n_location == "7") 
```
In 2002, the states that were observed at 7 locations are `CT`, `FL` and `NC`.

**Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**
```{r}
  group_by(tidy_brfss,locationabbr,year) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  ggplot(aes(y = n_location, x = year)) +
  geom_line(aes(color = locationabbr), alpha = .5) +
  labs(
    title = "Spaghetti plot",
    x = "Year",
    y = "Number of observations in each state"
  ) 
  
```

**Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**
```{r}
filter(tidy_brfss, year == "2002" | year == "2006" | year == "2010") %>% 
  filter(locationabbr == "NY") %>% 
  filter(response == "Excellent") %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  group_by(year) %>% 
  summarize(
    mean_proportion = mean(excellent),
    sd_proportion = sd(excellent)
    ) 
```

**For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**
```{r}
  group_by(tidy_brfss, year, locationabbr,response) %>% 
  summarize(mean = mean(data_value, na.rm = TRUE)) %>% 
  
  ggplot(aes(x = year, y = mean, color = response)) +
  geom_point() +
  facet_grid(.~response) +
    labs(
      title = "Average proportion in each response category",
      x = "Year",
      y = "Mean value of proportion for each response"
    ) +
  theme(legend.position = "bottom")
  
```

# Question 2
## Cleaning data

```{r}
cart_data = instacart
```
There are `r nrow(cart_data)` rows and `r ncol(cart_data)` coloums.

**How many aisles are there, and which aisles are the most items ordered from?**
```{r}
distinct(cart_data, aisle_id) %>% 
  count()

cart_data %>% 
  group_by(aisle) %>% 
  summarize(number = n()) %>% 
  arrange(desc(number))
```
There are `r distinct(cart_data, aisle_id) %>% count()` aisles. The most items ordered from is `fresh vegetables` and the second popular order item is `fresh fruits`.

**Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.**
```{r}
cart_data %>% 
  group_by(aisle) %>% 
  summarize(number = n()) %>% 
  mutate(aisle = fct_reorder(aisle, desc(number))) %>% 
  ggplot(aes(x = aisle, y = number)) +
  geom_point() +
  labs(
      title = "Number of items ordered in each aisle",
      x = "Items",
      y = "Number of items"
    ) +
  scale_y_log10(breaks = c(200000,20000, 2000, 200), 
                     labels = c("200000","20000", "2000", "200"),
                     limits = c(200, 200000)) +
  theme(legend.position = "bottom",
      axis.text.x = element_text(angle = 90, hjust = 1, size = rel(1)))
```

**Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.**
```{r}
popular_items = filter(cart_data, aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
group_by(aisle, product_name) %>% 
  summarize(number = n()) 

 top_baking_ingredients = filter(popular_items, aisle == "baking ingredients") %>% 
  arrange(desc(number)) %>%
    head(3) 
 
  top_dog_food_care = filter(popular_items, aisle == "dog food care") %>% 
    arrange(desc(number)) %>% 
    head(3)
  
  top_packaged_vegetables_fruits = filter(popular_items, aisle == "packaged vegetables fruits") %>% 
    arrange(desc(number)) %>% 
    head(3)
  
 popular_items = bind_rows(top_baking_ingredients,top_dog_food_care,top_packaged_vegetables_fruits)
```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)**
```{r}
mutate(cart_data, order_dow = factor(order_dow, labels  = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>% 
filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour)
```

# Question 3
```{r}
ny_noaa 
```

**Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?**
```{r}
tidy_ny_noaa = ny_noaa %>% 
  separate( date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.name[month],
    day = as.integer(day),
    prcp = prcp/10,
    snow = as.integer(snow),
    snwd = as.integer(snwd),
    tmax = as.numeric(tmax)/10,
    tmin = as.numeric(tmin)/10
    )

filter(tidy_ny_noaa, snow != "NA") %>% 
  group_by(snow) %>% 
  summarize(number = n()) %>% 
  arrange(desc(number)) %>% 
  head(4)
```
So, `0` is the most commonly ovserved value. Because in most days of a year, New York would not snow.

**Make a two-panel plot showing the average MAX temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?**
```{r}
filtered_ny_noaa = filter(tidy_ny_noaa, month %in% c("January" , "July")) %>% 
group_by(month, year, id) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  filter(mean_tmax != "NaN") 
  
ggplot(filtered_ny_noaa, aes(x = year, y = mean_tmax, group = year)) +
  geom_boxplot() +
  facet_grid(~month) +
  labs(
    x = "Year",
    y = "Average maximum temperature",
    title = "Average maximum temperature across years in January and July"
  )
```

**Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.**
```{r}
temperature_plot = 
tidy_ny_noaa %>% 
  filter(tmax != "NA", tmin != "NA") %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  labs(
    x = "Minimum temperature (ºC)",
    y = "Maximum temperature (ºC)",
    title = "Maximum vs Minimum Temperature"
  )

snowfall_plot = 
  tidy_ny_noaa %>% 
  filter(snow > 0 , snow < 100) %>% 
  ggplot(aes(x = year, y = snow)) +
  geom_boxplot(aes(group = year))+
  labs(
    x = "Year",
    y = "Snowfall (mm)",
    title = "Distribution of snowfall values over years"
  )

temperature_plot + snowfall_plot
```


